Two-component thermal dust emission model: application to the {\it PLANCK} HFI maps

### 2014-07-09 ms.tex
a bunch of clean up in reddening residual discussion

@@

We calibrate to reddening empirically rather than derive
a relationship between $\tau_{545}$ and reddening by introducing additional 
assumptions about the dust grain physics and size distribution. To achieve this
empirical calibration, we must adopt a set of calibrator objects for which true
optical reddening is known. There are various possibilities at our disposal. 
\cite{planckdust} calibrated their radiance and $\tau_{353}$ maps to $E(B-V)$ 
using broadband Sloan Digital Sky Survey \citep[SDSS;][]{sdss} photometry for a
set of $\sim$10$^5$ quasars. The SFD calibration was originally tied to a 
sample of 384 elliptical galaxies, but was later revised by \cite{schlafly11} 
using $\sim$250,000 stars with both spectroscopy and broadband photometry 
available from SDSS. To calibrate our $\tau_{545}$ to reddening, we make use of
the same set of stars studied by \cite{schlafly11}.

@@

Given a library of stellar atmosphere models, the spectral lines of these stars
can be used to predict their intrinsic optical broadband colors. The `true' 
reddening is then simply the difference in the observed $g-r$ color and the 
$g-r$ color predicted from the spectral lines. Applying a color transformation 
then yields `true' $E(B-V)$ values for $\sim$260,000 lines of sight. Throughout
our SSPP calibration analysis, we restrict to the $\sim$230,000 lines of sight 
with $|b|$$>$20$^{\circ}$ in order to avoid stars which may not lie behind the 
full dust column. The calibration of two-component optical depth to $E(B-V)$ is
performed as a linear regression of $E(B-V)_{SSPP}$ versus $\tau_{545}$. 
$\tau_{545}$ is considered to be the independent variable in this regression, 
as we ultimately wish to predict $E(B-V)$ as a function of optical depth, and 
$\tau_{545}$ has much higher S/N than the SSPP $E(B-V)$ estimates.

@@

This regression is illustrated in Figure \ref{fig:calib}. As expected, there 
appears to be a strong linear correlation between $E(B-V)_{SSPP}$ and 
$\tau_{545}$. The conversion factor from $\tau_{545}$ to $E(B-V)$ is 
2.46$\times$10$^{3}$. Reassuringly, the best-fit offset is quite close to zero,
$\sim$0.6 mmag.

@@

All four residual plots in Figure 12 show systematic problems at some level. 
The most striking systematic trend is the `bending' behavior of the reddening
residuals versus $E(B-V)_{SFD}$ (top panels), with the median residual 
bottoming out near $-10$ mmag around $E(B-V)_{SFD}$=0.15 mag. This behavior is 
common to both $E(B-V)_{2comp}$ and $E(B-V)_{mbb}$, and in fact was first noted
in the residuals of $E(B-V)_{SFD}$ itself relative to $E(B-V)_{SSPP}$ by SF11 
(see their Figure 6). Such a bending behavior is troubling because it could 
indicate a nonlinearity common to many FIR reddening predictions based on
column densities inferred from dust emission. Alternatively, because the SF11 
stars are distributed over the sky in a highly non-uniform manner, the bend 
could arise from aliasing of discrepancies particular to certain sky regions 
(e.g. inner vs. outer Galaxy) on to the $E(B-V)_{SFD}$ axis.

@@

The obvious culprit for any potential nonlinearity in FIR-based reddening
estimates is a faulty temperature correction. For this reason, we have included
the bottom panels of Figure \ref{fig:resid}, to check for the presence of
a temperature dependence of the reddening residuals. Indeed, in both the
two-component and single-MBB cases there exists some systematic dependence of
the reddening residuals on temperature. For $T_{mbb}$$\gtrsim$19 K, the 
median residual is reasonably flat, but at lower temperatures (the lowest 
temperature $\sim$20\% of the sky), the median shows trends at the $\sim$10 
mmag level. On the other hand, the median residual in the two-component case 
shows a clear trend of decreasing with increasing $T_2$ over the entire range
of $T_2$ shown, with a peak-to-peak amplitude of $\sim$20 mmag.

@@

The following is a list of dust model modifications we tested, but 
proved to have little impact on the reddening residual trends versus either 
$E(B-V)_{SFD}$ or $T_2$:

@@

However, we found that changing the zero level offsets of the input maps 
had a significant effect on the strength of the anticorrelation between median 
reddening residual and $T_2$. In particular, we experimented with perturbing 
the zero level offset of \PLANCK~857 GHz while correspondingly changing the 
zero levels of the remaining \PLANCK~maps based on the prescription of 
$\S$\ref{sec:relzero}. We also experimented with changing the zero level of SFD
\verb|i100|, independent of the other zero levels. Unfortunately, completely
flattening the reddening residual dependence on $T_2$ required unreasonably 
large zero level changes. For example, flattening the $T_2$ residual required
adding $\gtrsim$0.6 MJy/sr to the \verb|i100| map. Such an offset is 
implausible, being an order of magnitude larger than the nominal \verb|i100| 
zero level uncertainty quoted by SFD98, and comparable to the entire 3000 GHz 
CIB monopole signal. Furthermore, we note that even these large zero 
level modifications had virtually no effect in eliminating the reddening 
residual `bend' versus $E(B-V)_{SFD}$. Thus, changing the zero level 
offsets showed hints of promise in rectifying the residual temperature 
dependence, but cannot completely resolve the systematic trends in reddening 
residuals.
% last few words here are kind of bad, need better terminology

The ony solution we have been able to identify that both removes the `bend' vs.
$E(B-V)_{SFD}$ and the temperature dependence of the residual is cutting 
out the ecliptic plane, restricting to $|\beta|>20^{\circ}$ . In this case, we 
completely eliminated the bending behavior of the residual versus 
$E(B-V)_{SFD}$, and significantly reduced the $T_2$ dependence (see Figure
\ref{fig:resid_ecl}), to a peak-to-peak amplitude of only $\sim$10 mmag. We
still included the single-MBB plots (right column), to show that 
the bend versus $E(B-V)_{SFD}$ is eliminated by the $|\beta|$ cut, even
for the single-MBB model. However, the single-MBB residuals still differ
systematically from zero for $T$$\lesssim$$19$ K. Perhaps these improvements in
the two-component reddening residuals after restricting to high ecliptic 
latitude should come as no surprise, given that the ecliptic plane is the most 
obvious systematic problem with our temperature map upon inspecting the 
full-sky results shown in Figure \ref{fig:results}.

@@

After cutting the ecliptic plane, we found that only small zero level
perturbations were required to fully flatten the temperature residuals,
while still maintaining flat residuals versus $E(B-V)_{SFD}$. The optimal
offsets we found were $\pm$0.08 MJy/sr to \verb|i100| and 857 GHz respectively
(see Figure \ref{fig:resid_offs}). These offsets are well within reason, 
given the nominal zero level uncertainties quoted in Table QQ.

>>>>

% mention that this is especially true because Planck gives such a huge
% increase in temperature resolution relative to SFD (and possibly reference
% temperature comparison figure as well)

@@

We calibrate to reddening empirically rather than derive
a relationship between $\tau_{545}$ and reddening by introducing additional 
assumptions about the dust grain physics and size distribution. To achieve this
empirical calibration, we must adopt a set of calibrator objects for which true
optical reddening is known. There are various possibilities at our disposal. 
\cite{planckdust} calibrated their radiance and $\tau_{353}$ maps to $E(B-V)$ 
using broadband Sloan Digital Sky Survey \citep[SDSS;][]{sdss} photometry for a
set of $\sim$10$^5$ quasars. The SFD calibration was originally tied to a 
sample of 384 elliptical galaxies, but was later revised by 
\citet[hereafter SF11]{schlafly11} based on $\sim$260,000 stars with both 
spectroscopy and broadband photometry available from SDSS.

@@

To calibrate our two-component optical depth to reddening, we make use of the 
stellar sample from SF11. Given a library of model stellar atmospheres, the 
spectral lines of these stars can be used to predict their intrinsic optical 
broadband colors. The `true' reddening is then simply the difference between 
the observed $g-r$ color and the $g-r$ color predicted from the spectral lines.
Applying a color transformation then yields `true' $E(B-V)$ values for 
$\sim$260,000 lines of sight. Throughout our SSPP calibration analysis, we 
restrict to the $\sim$230,000 lines of sight with $|b|$$>$20$^{\circ}$ in order
to avoid stars which may not lie behind the full dust column. The calibration 
of two-component optical depth to $E(B-V)$ is performed as a linear regression 
of $E(B-V)_{SSPP}$ versus $\tau_{545}$. $\tau_{545}$ is considered to be the 
independent variable in this regression, as we ultimately wish to predict 
$E(B-V)$ as a function of optical depth, and $\tau_{545}$ has much higher S/N 
than the SSPP $E(B-V)$ estimates.

@@

This regression is illustrated in Figure \ref{fig:calib}. As expected, there 
is a strong linear correlation between $E(B-V)_{SSPP}$ and $\tau_{545}$. The 
conversion factor from $\tau_{545}$ to $E(B-V)$ is 2.46$\times$10$^{3}$. 
Reassuringly, the best-fit offset is quite close to zero, $\sim$0.6 mmag.

@@

All four residual plots in Figure 12 show systematic problems at some level. 
The most striking systematic trend is the `bending' behavior of the reddening
residuals versus $E(B-V)_{SFD}$ (top panels), with the median residual 
bottoming out near $-10$ mmag at $E(B-V)_{SFD}$$\approx$0.15 mag. This behavior
is common to both $E(B-V)_{2comp}$ and $E(B-V)_{mbb}$, and in fact was first 
noted in the residuals of $E(B-V)_{SFD}$ itself relative to $E(B-V)_{SSPP}$ by 
SF11 (see their Figure 6). Such a bending behavior is troubling because it 
could indicate a nonlinearity common to many FIR reddening predictions based on
column densities inferred from dust emission. Alternatively, because the SF11 
stars are distributed over the sky in a highly non-uniform manner, the bend 
could arise from aliasing of discrepancies particular to certain sky regions 
(e.g. inner vs. outer Galaxy) on to the $E(B-V)_{SFD}$ axis.

@@

The obvious culprit for any potential nonlinearity in FIR-based reddening
estimates is a faulty temperature correction. For this reason, we have included
the bottom panels of Figure \ref{fig:resid}, to check for the presence of
a temperature dependence of the reddening residuals. Indeed, in both the
two-component and single-MBB cases there exists some systematic dependence of
the reddening residuals on temperature. For $T_{mbb}$$\gtrsim$19 K, the 
median residual is reasonably flat, but at lower temperatures (the lowest 
temperature $\sim$20\% of the sky), the median shows trends at the $\sim$10 
mmag level. On the other hand, the median residual in the two-component case 
trends downward with increasing $T_2$ over the entire $T_2$ range shown, with a
peak-to-peak amplitude of $\sim$20 mmag.

@@

The following is a list of dust model modifications we tested, but which
proved to have little impact on the reddening residual trends as a function 
of either $E(B-V)_{SFD}$ or $T_2$:

@@

However, we found that changing the zero level offsets of the input maps 
had a significant effect on the strength of the anticorrelation between median 
reddening residual and $T_2$. In particular, we experimented with perturbing 
the zero level offset of \PLANCK~857 GHz while correspondingly changing the 
zero levels of the remaining \PLANCK~maps based on the prescription of 
$\S$\ref{sec:relzero}. We also experimented with changing the zero level of SFD
\verb|i100|, independent of the other zero levels. Unfortunately, completely
flattening the reddening residual dependence on $T_2$ required unreasonably 
large zero level modifications. For example, flattening the $T_2$ residual 
required adding $\gtrsim$0.6 MJy/sr to the \verb|i100| map. Such an offset is 
implausible, being an order of magnitude larger than the nominal \verb|i100| 
zero level uncertainty quoted by SFD98, and comparable to the entire 3000 GHz 
CIB monopole signal. Furthermore, we note that even these large zero 
level modifications had virtually no effect in eliminating the reddening 
residual `bend' versus $E(B-V)_{SFD}$. Thus, changing the zero level 
offsets showed hints of promise in rectifying the reddening residual 
temperature dependence, but cannot completely resolve the systematic trends in 
reddening residuals.
% last few words here are kind of bad, need better terminology

The only solution we have been able to identify that both removes the `bend' 
vs. $E(B-V)_{SFD}$ and simultaneously reduces the temperature dependence of the
reddening residuals is cutting out the ecliptic plane by restricting to 
$|\beta|>20^{\circ}$. In this case, we completely eliminated the bending 
behavior of the residual versus $E(B-V)_{SFD}$, and significantly reduced the 
$T_2$ dependence to a peak-to-peak amplitude of only $\sim$10 mmag (see Figure 
\ref{fig:resid_ecl}). Figure \ref{fig:resid_ecl} still includes the single-MBB 
plots (right column), to show that the bend versus $E(B-V)_{SFD}$ is eliminated
by the $|\beta|$ cut, even for the single-MBB model. However, the single-MBB 
residuals still differ systematically from zero for $T$$\lesssim$$19$ K. 
Perhaps the improvements in the two-component reddening residuals after 
restricting to high ecliptic latitude should come as no surprise, given that 
the ecliptic plane is the most obvious systematic problem with our temperature 
map (see the full-sky results shown in Figure \ref{fig:results}).

@@

After cutting the ecliptic plane, we found that only small zero level
perturbations were required to fully flatten the temperature residuals,
while still maintaining flat residuals versus $E(B-V)_{SFD}$. The optimal
offsets we found were $\pm$0.08 MJy/sr to \verb|i100| and 857 GHz respectively
(see Figure \ref{fig:resid_offs}). These offsets are well within reason, 
given the nominal zero level uncertainties quoted in Table \ref{table:offs}.
